Eddy covariance towers as sentinels of abnormal radioactive material releases

### 2024-11-07 manuscript/manuscript.tex
stash text edits

In order to flag event detections, we defined several adjustable hyperparameters to account for uncertainty surrounding atmospheric transport and the timing of any possible responses to material deposition. The first hyperparameter we defined, \emph{wind\_tolerance}, reflects the range around the bearing from the facility to the tower location which we count as towards the tower. In the base default case, we set \emph{wind\_tolerance} to 10 degrees, but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees. The second hyperparameter we defined, \emph{event\_quantile\_effect}, handles the unknown travel time of any materials to be deposited via an abnormal release or what, if any, delays exist between deposition and vegetation response. Rather than simply selecting the maximum value of $\beta$3 during a given window or the value at the exact time of the event, the value of \emph{event\_quantile\_effect} is set to correspond with a specific quantile of all $\beta$3 values during the event window. In the base default case, this was set at 0.9, but we explored values as low as 0.5. All statistical analyses were carried out using the statsmodels Python package (Seabold and Perktold, 2009). Our processing scripts are openly available at [doi link], and we refer readers to the original providers for data access.

@@

Despite the similar effect size at the three tower sites during the abnormal release event, they differed substantially in the degree to which they uniquely flagged the known release event. At the BE-Lon and BE-Vie sites, our event detection algorithm identified the known release and only a few other time periods as potential abnormal releases (Figure 3). Conversely, at the northerly BE-Bra site, the algorithm was not able to identify the known release event distinct from background noise. This can be seen in the rate of event detections ($\alpha$) at each site, whereby it was less than 1\% at the BE-Lon and BE-Vie sites, but was greater than 2\% at the BE-Bra site. It is notable that this ability to uniquely detect the known release along with a plausible number of additional event detections corresponds with the fraction of the abnormal release period when the wind was pointed from the release location towards the respective tower (using the default base case wind tolerance of 10 deg) which was much lower at the BE-Bra site (5\%) compared to the BE-Lon and BE-Vie sites (34\% Table 1).

To increase our confidence that event detections were not dependent on the specific hyperparameters settings in our base-case setup, we ran an exhaustive cross-validation set of hyperparameter experiments. For this effort, we tested every possible combination of values for wind tolerance, number of days, and the event quantile effect. Across reasonable values of specific hyperparameters, we found that their magnitudes had little effect on the rate at which events were flagged as potential abnormal releases as even in the most extreme cases, we did not see detection rates exceeding 1\%. We did, however, observe that the event detection was related to the window length and event quantile parameters as shown by the slope of the line in Figure 4. Although increasing values of the event quantile parameter would hypothetically increase event detection rates, we do not see value in setting this parameter below the median. Similarly, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not see value in setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.

>>>>

In order to flag event detections, we defined several adjustable hyperparameters to account for uncertainty surrounding atmospheric transport and the timing of any possible responses to material deposition. The first hyperparameter we defined, \emph{wind\_tolerance}, reflects the range around the bearing from the facility to the tower location which we count as towards the tower. In the base default case, we set \emph{wind\_tolerance} to 10 degrees, but we explored setting it to smaller values down to 5 degrees and larger values up to 20 degrees. The second hyperparameter we defined, \emph{event\_quantile\_effect}, handles the unknown travel time of any materials to be deposited via an abnormal release or what, if any, delays exist between deposition and vegetation response. Rather than simply selecting the maximum value of $\beta$3 during a given window or the value at the exact time of the event, the value of \emph{event\_quantile\_effect} is set to correspond with a specific quantile of all $\beta$3 values during the event window. In the base default case, this was set at 0.9, but we explored values as low as 0.5. Taken together, here is the overall logic we used to flag an event detection: consider a scenario where the effect size value (corresponding to the \emph{event\_quantile\_effect} of the distribution of effect sizes during the \emph{n\_days} window length) during a known event is X at a tower which is Z bearing from the known release location. An event detection would be flagged during any period where the effect size is $\geq$ X and the wind direction is Z $\pm$ \emph{wind\_tolerance}.

All statistical analyses in the present study were carried out using the statsmodels Python package (Seabold and Perktold, 2009). Our processing scripts are openly available at \href{https://doi.org/10.5281/zenodo.13845254}{https://doi.org/10.5281/zenodo.13845254}, and we refer readers to the original providers for data access.

@@

Despite the similar effect size at the three tower sites during the abnormal release event, they differed substantially in the degree to which they uniquely flagged the known release event. At the BE-Lon and BE-Vie sites, our event detection algorithm identified the known release and only a few other time periods as potential abnormal releases (Figure 3). Conversely, at the northerly BE-Bra site, the algorithm was not able to identify the known release event distinct from background noise. This can be seen in the rate of event detections ($\alpha$) at each site, calculated as the number of event detections divided by the total number of rolling windows in the period of record. This rate was less than 1\% at the BE-Lon and BE-Vie sites, but was greater than 2\% at the BE-Bra site. It is notable that this ability to uniquely detect the known release along with a plausible number of additional event detections corresponds with the fraction of the abnormal release period when the wind was pointed from the release location towards the respective tower. The fraction of the period of record where the wind was pointed from the facility toward the flux towers was much lower at the BE-Bra site (5\%) compared to the BE-Lon and BE-Vie sites (34\% Table 1).

To increase our confidence that event detections were not dependent on the specific hyperparameters settings in our base-case setup, we ran an exhaustive cross-validation set of hyperparameter experiments. For this effort, we tested every possible combination of values for wind tolerance, number of days, and the event quantile effect. Across reasonable values of specific hyperparameters, we found that their magnitudes had little effect on the rate at which events were flagged as potential abnormal releases as even in the most extreme cases, we did not see detection rates exceeding 1\%. We did, however, observe that the event detection was related to the window length and event quantile parameters as shown by the slope of the line in Figure 4. Although increasing values of the event quantile parameter above the limit of our tests ($>$ 0.9) would hypothetically increase event detection rates, we do not show this due to the noisiness of the extreme quantiles of effect size distributions. Additionally, we do not show tests setting this parameter below the median because it would lead to  weak sensitivity to abnormal events. In a similar fashion as the event quantile parameter, increasing the magnitude of the window length parameter would also likely increase the event detection rate but here too we do not show tests setting this parameter beyond 14 days given the results of prior atmospheric modeling efforts showing limited deposition at such long time scales \citep[e.g.][]{meszarosPredictabilityDispersionFukushimaderived2016}.
